Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-24 src/ms.tex
referee comments

\newcommand{\taucz}{$\tau_\mathrm{cz}$\xspace}
\newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
%\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~\mathrm{Ro_\odot}$\xspace}
\newcommand{\lamostkep}{LAMOST--Kepler\xspace}
\newcommand{\lamostmcq}{LAMOST--McQuillan\xspace}
\newcommand{\lamostsan}{LAMOST--Santos\xspace}
\newcommand{\teffmin}{5500~K\xspace}
\newcommand{\agerange}{$\sim$2--6~Gyr\xspace}
\newcommand{\agemin}{$\sim$2--3~Gyr\xspace}

@@

\subsection{The Sun}
To place the Sun in the context of the long-period pile-up, we use up-to-date estimates of the Sun's effective temperature, rotation period, and age. Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K. For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d. \edit1{While differential rotation studies have inferred polar rotation periods of $\approx$~36~d from helioseismology \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}. We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}. \edit1{For the color of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.82$ \citep{Casagrande2018}.}

@@

It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability. This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values. Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
    \label{fig:gaia-mcquillan}
\end{figure}

>>>>

\newcommand{\taucz}{$\tau_\mathrm{cz}$\xspace}
\newcommand{\rocrit}{$\mathrm{Ro_{crit}}$\xspace}
\newcommand{\rothresh}{$\mathrm{Ro_{thresh}}$\xspace}
%\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~2$\xspace}
\newcommand{\rocritfinal}{$\mathrm{Ro_{crit}}~\lesssim~\mathrm{Ro_\odot}$\xspace}
\newcommand{\lamostkep}{LAMOST--Kepler\xspace}
\newcommand{\lamostmcq}{LAMOST--McQuillan\xspace}
\newcommand{\lamostsan}{LAMOST--Santos\xspace}
\newcommand{\teffmin}{5500~K\xspace}
\newcommand{\agerange}{$\sim$2--6~Gyr\xspace}
\newcommand{\agemin}{$\sim$2--3~Gyr\xspace}

@@

\subsection{The Sun}
To place the Sun in the context of the long-period pile-up, we use up-to-date estimates of the Sun's effective temperature, rotation period, and age. Following the IAU 2015 Resolution B3 \citep{Prsa2016}, we take the nominal effective temperature of the Sun to be $\mathcal{T}^\mathrm{N}_\mathrm{eff,\odot} = 5772$~K. For the solar rotation period we adopt the equatorial rotation period of $P_\mathrm{eq,\odot}~\approx~25.4$~d. \edit1{While differential rotation studies have inferred polar rotation periods of $\approx$~36~d from helioseismology \citep[][and references therein]{Thompson2003}, sunspots are rarely observed outside the activity belts, roughly $\pm 30^\circ$ from the equator \citep[][and references therein]{Hathaway2015}. We therefore adopt 27~d as an approximate upper limit to the rotation period of the Sun, as it would be observed by Kepler through rotational brightness modulations.} The age of the Sun is assumed to be 4.567~Gyr from Pb-Pb dating of calcium-aluminum inclusions and chondrules recovered from primitive meteorites \citep[][and references therein]{Bahcall1995}. \edit1{For the color and apparent magnitude of the Sun in the Gaia bandpasses we adopt $G_\mathrm{BP}-G_\mathrm{RP} = 0.818$~mag and $G = -26.895$~mag \citep{Casagrande2018}.}

@@

It is evident from Figure~\ref{fig:gap} that the long-period pile-up also coincides with a steep gradient in variability amplitude, such that a hidden population of pile-up stars may lie just beyond the edge of detectability. This effect, in addition to the \teff offsets mentioned above, can further bias our inference of \rocrit to lower values. Regardless of the biases mentioned above and the true value of \rocrit, the sample considered clearly indicates that the long-period pile-up is apparent for stars with \teff~$\gtrsim$~\teffmin with Rossby numbers lower than those associated with a detection edge. \edit1{In other words, the fact that we observe a pile-up indicates that if WMB is assumed to be the cause, then \rocrit~$\lesssim$~\rothresh, where \rothresh again is assumed to be a detection threshold.}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}. The grey lines indicate empirical cluster sequences derived in \citet{Curtis2020}. From top to bottom, the cluster sequences are Pleiades (0.12~Gyr), Praesepe (0.67~Gyr), NGC 6811 (1~Gyr), and NGC 6819 + Ruprecht 147 (2.5~Gyr).}
    \label{fig:gaia-mcquillan}
\end{figure}
